---
title: "Metan analysis"
author: "Hardlife Chinwa"
date: "2024-08-16"
output: html_document
---
Getting started
In this section, we will use the data in data_ge. For more information, please, see ?data_ge. Other data sets can be used provided that the following columns are in the dataset: environment, genotype, block/replicate and response variable(s). See the section Rendering engine to know how HTML tables were generated.

The first step is to inspect the data with the function inspect()

```{r}
yld=read.csv("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/DPhil/RGB for indirect selection/Field trial data final/RGB_data_final/Goundnut_yield_metan.csv",header=TRUE)
```

```{r}
attach(yld)
```

```{r}
yld$Environment=as.factor(yld$Environment)
yld$Replication=as.factor(yld$Replication)
yld$Genotype=as.factor(yld$Genotype)
```

```{r}
str(yld)
```

```{r}
library(metan)
library(plyr)
inspect <- 
inspect(yld,
        verbose = FALSE,
        plot = TRUE)
```

```{r}
inspect
```

Analysis of single experiments using mixed-models
The function gamem() may be used to analyze single experiments (one-way experiments) using a mixed-effect model according to the following model:

yij=μ+αi+τj+εij
where yij
 is the value observed for the ith genotype in the jth replicate (i = 1, 2, … g; j = 1, 2, .., r); being g and r the number of genotypes and replicates, respectively; αi
 is the random effect of the ith genotype; τj
 is the fixed effect of the jth replicate; and εij
 is the random error associated to yij
. In this example, we will use the example data data_g from metan package.
```{r}
gen_mod <- gamem(yld, Genotype, Environment,
                 resp = c(SHW,SP,PY,Biomass,HI,GY,Severity,GGA,Necrosis,a,u,Chlorosis,b,v,CSI))
```

The easiest way of obtaining the results of the model above is by using the function get_model_data(). Let’s do it.

Details of the analysis

```{r}
get_model_data(gen_mod, "details")
# Class of the model: gamem
# Variable extracted: details
```

Likelihood ratio test for genotype effect

```{r}
get_model_data(gen_mod, "lrt")
# Class of the model: gamem
# Variable extracted: lrt
```

Variance components and genetic parameters

```{r}
mod=get_model_data(gen_mod, "genpar")
# Class of the model: gamem
# Variable extracted: genpar
mod
```

```{r}
variances=write.csv(mod,"Genetic variances of groundnut genotypes.csv")
```
#####predicted means
```{r}
BLUP_predict=get_model_data(gen_mod, "blupg")
# Class of the model: gamem
# Variable extracted: blupg
```
```{r}
write.csv(BLUP_predict,"Predicted BLUP means.csv")
```

In the above example, the experimental design was a complete randomized block. It is also possible to analyze an experiment conducted in an alpha-lattice design with the function gamem(). In this case, the following model is fitted:

yijk=μ+αi+γj+(γτ)jk+εijk

where yijk
 is the observed value of the ith genotype in the kth block of the jth replicate (i = 1, 2, … g; j = 1, 2, .., r; k = 1, 2, .., b); respectively; αi
 is the random effect of the ith genotype; γj
 is the fixed effect of the jth complete replicate; (γτ)jk
 is the random effect of the kth incomplete block nested within the j replicate; and εijk
 is the random error associated to yijk
. In this example, we will use the example data data_alpha from metan package.
###for alpha lattice

```{r}
#en_alpha <- gamem(data_alpha, Genotype, Farmer, Actual_yield.kg.ha, block = Farmer)
```


all variables with significance
```{r}
#get_model_data(gen_alpha, "lrt")
```

The BLUP model for MET trials
The simplest and well-known linear model with interaction effect used to analyze data from multi-environment trials is
yijk=μ+αi+τj+(ατ)ij+γjk+εijk

where yijk
 is the response variable (e.g., grain yield) observed in the kth block of the ith genotype in the jth environment (i = 1, 2, …, g; j = 1, 2, …, e; k = 1, 2, …, b); μ
 is the grand mean; αi
 is the effect of the ith genotype; τj
 is the effect of the jth environment; (ατ)ij
 is the interaction effect of the ith genotype with the jth environment; γjk
 is the effect of the kth block within the jth environment; and εijk
 is the random error. In a mixed-effect model assuming αi
 and (ατ)ij
 to be random effects, the above model can be rewritten as follows

y=Xβ+Zu+ε

where y is an n[=∑ej=1(gb)]×1
 vector of response variable y=[y111,y112,…,ygeb]′
; β
 is an (eb)×1
 vector of unknown fixed effects β=[γ11,γ12,...,γeb]′
; u is an m[=g+ge]×1
 vector of random effects u=[α1,α2,...,αg,(ατ)11,(ατ)12,...,(ατ)ge]′
; X is an n×(eb)
 design matrix relating y to β
; Z is an n×m
 design matrix relating y to u ; ε
 is an n×1
 vector of random errors ε=[y111,y112,…,ygeb]′
;

The vectors β
 and u are estimated using the well-known mixed model equation (Henderson 1975).
 
 [β^u^]=[X′R−1XZ′R−1XX′R−1ZZ′R−1Z+G−1]−[X′R−1yZ′R−1y]
 
where G and R are the variance-covariance matrices for random-effect vector u and residual vector ε
, respectively.

The function gamem_met() is used to fit the linear mixed-effect model. The first argument is the data, in our example data_ge. The arguments (env, gen, and rep) are the name of the columns that contains the levels for environments, genotypes, and replications, respectively.The argument (resp) is the response variable to be analyzed. The function allow a single variable (in this case GY) or a vector of response variables. Here, we will use everything() to analyse all numeric variables in the data. By default, genotype and genotype-vs-environment interaction are assumed to be random effects. Other effects may be considered using the random argument. The last argument (verbose) control if the code is run silently or not.

```{r}
str(yld)
```

```{r}
yld$Environment=as.factor(yld$Environment)
yld$Farmer=as.factor(yld$Farmer)
yld$Genotype=as.factor(yld$Genotype)
```

```{r}
yld$Environment=as.factor(yld$Environment)
yld$Replication=as.factor(yld$Replication)
yld$Genotype=as.factor(yld$Genotype)
```


```{r}
mixed_mod_nec <- 
  gamem_met(yld,
            env = Environment,
            gen = Genotype,
            rep = Replication,
            resp = c(Necrosis,Chlorosis,SP,PY,Biomass,HI,SHW,GY,GGA,Chlorosis,a,u,Severity,b,v,CSI),
            random = "gen",            #Default
            by=NULL,
            verbose = FALSE) #Default
mixed_mod_nec
```

Diagnostic plot for residuals
The S3 generic function plot() is used to generate diagnostic plots of residuals of the model

```{r}
plot(mixed_mod_chl)
# `geom_smooth()` using formula = 'y ~ x'
# `geom_smooth()` using formula = 'y ~ x'
```

The normality of the random effects of genotype and interaction effects may be also obtained by using type = "re".

```{r}
plot(mixed_mod_chl, type = "re")
```
```{r}
plot(mixed_mod_chl, type = "re")
```

Printing the model outputs
Likelihood Ratio Tests
The output LRT contains the Likelihood Ratio Tests for genotype and genotype-vs-environment random effects. We can get these values with get_model_data()

```{r}
data <- get_model_data(mixed_mod_chl, "lrt")
data
```

Variance components and genetic parameters
In the output ESTIMATES, beyond the variance components for the declared random effects, some important parameters are also shown. Heribatility is the broad-sense heritability, h2g
, estimated by
h2g=σ^2gσ^2g+σ^2i+σ^2e

where σ^2g
 is the genotypic variance; σ^2i
 is the genotype-by-environment interaction variance; and σ^2e
 is the residual variance.

GEIr2 is the coefficient of determination of the interaction effects, r2i
, estimated by

r2i=σ^2iσ^2g+σ^2i+σ^2e
Heribatility of means is the heribability on the mean basis, h2gm
, estimated by

h2gm=σ^2g[σ^2g+σ^2i/e+σ^2e/(eb)]

where e and b are the number of environments and blocks, respectively; Accuracy is the accuracy of selection, Ac, estimated by
Ac=h2gm−−−√

rge is the genotype-environment correlation, rge
, estimated by

rge=σ^2gσ^2g+σ^2i

CVg and CVr are the the genotypic coefficient of variation and the residual coefficient of variation estimated, respectively, by
CVg=(σ^2g−−√/μ)×100
and
CVr=(σ^2e−−√/μ)×100
where μ
 is the grand mean.

CV ratio is the ratio between genotypic and residual coefficient of variation.

```{r}
data <- get_model_data(mixed_mod_chl)
# Class of the model: waasb
# Variable extracted: genpar
data
```

BLUP for genotypes

```{r}
BLUP_rank_chl=mixed_mod_sev$Severity$BLUPgen
```

```{r}
BLUP_rank_sev
```

```{r}
write.csv(BLUP_rank_sev,"Ranked BLUP means on disease.csv")
```

The function get_model_data() may be used to easily get the data from a model fitted with the function gamem_met(), especially when more than one variables are used. The following code return the predicted mean of each genotype for five variables of the data data_ge2

```{r}
BLUP_plotted=get_model_data(mixed_mod_nec, what = "blupg")
```

```{r}
write.csv(BLUP_plotted,"BLUP means used to plot the means_sev.csv")
```

Plotting the BLUP for genotypes

```{r}
library(ggplot2)
a <- plot_blup(mixed_mod_nec)
b <- plot_blup(mixed_mod_nec, 
               col.shape  =  c("gray20", "gray80"),
               plot_theme = theme_metan(grid = "y")) +
  coord_flip()
arrange_ggplot(a, b, tag_levels = "a")
plot(a)
```

This output shows the predicted means for genotypes. BLUPg is the genotypic effect (g^i)
, which considering balanced data and genotype as random effect is estimated by

g^i=h2g(y¯i.−y¯..)

where h2g
 is the shrinkage effect for genotype. Predicted is the predicted mean estimated by
g^i+μ

where μ
 is the grand mean. LL and UL are the lower and upper limits, respectively, estimated by
(g^i+μ)±CI
with
CI=t×((1−Ac)×σ2g)−−−−−−−−−−−−−√

where t
 is the Student’s t value for a two-tailed t test at a given probability error; Ac
 is the accuracy of selection and σ2g
 is the genotypic variance.

BLUP for genotypes X environment combination

```{r}
BLUPGe=mixed_mod$GY$BLUPint
```

```{r}
write.csv(BLUPGe,"BLUP means for genotype by environment.csv")
```

This output shows the predicted means for each genotype and environment combination. BLUPg is the genotypic effect described above. BLUPge is the genotypic effect of the ith genotype in the jth environment (g^ij)
, which considering balanced data and genotype as random effect is estimated by
g^ij=h2g(y¯i.−y¯..)+h2ge(yij−y¯i.−y¯.j+y¯..)
where h2ge
 is the shrinkage effect for the genotype-by-environment interaction; BLUPg+ge is BLUPg+BLUPge
; Predicted is the predicted mean (y^ij
) estimated by
y^ij=y¯.j+BLUPg+ge

Some useful information
The following pieces of information are provided in Details output. Nenv, the number of environments in the analysis; Ngen the number of genotypes in the analysis; mresp The value attributed to the highest value of the response variable after rescaling it; wresp The weight of the response variable for estimating the WAASBY index. Mean the grand mean; SE the standard error of the mean; SD the standard deviation. CV the coefficient of variation of the phenotypic means, estimating WAASB, Min the minimum value observed (returning the genotype and environment), Max the maximum value observed (returning the genotype and environment); MinENV the environment with the lower mean, MaxENV the environment with the larger mean observed, MinGEN the genotype with the lower mean, MaxGEN the genotype with the larger.

```{r}
data <- get_model_data(mixed_mod, "details")
# Class of the model: waasb
# Variable extracted: details
data
```

The WAASB object
The function waasb() function computes the Weighted Average of the Absolute Scores considering all possible IPCA from the Singular Value Decomposition of the BLUPs for genotype-vs-environment interaction effects obtained by an Linear Mixed-effect Model (Olivoto et al. 2019), as follows:

WAASBi=∑k=1p|IPCAik×EPk|/∑k=1pEPk

where WAASBi
 is the weighted average of absolute scores of the ith genotype; IPCAik
 is the scores of the ith genotype in the kth IPCA; and EPk
 is the explained variance of the kth PCA for k=1,2,..,p
, p=min(g−1;e−1)
.

```{r}
waasb_model <- 
  waasb(yld,
        env = Environment,
        gen = Genotype,
        rep = Replication,
        resp = c(Severity,Chlorosis,SP,PY,Biomass,HI,SHW,Necrosis,GGA,Chlorosis,a,u,GY,b,v,CSI),
        random = "gen", #Default
        verbose = TRUE) #Default
```

```{r}
data <- waasb_model$GY$model
data
```

The output generated by the waasb() function is very similar to those generated by the waas() function. The main difference here, is that the singular value decomposition is based on the BLUP for GEI effects matrix.

Eigenvalues of the BLUP_GEI matrix

```{r}
data <- waasb_model$GY$PCA
data
```

```{r}
write.csv(data,"PCA for the model.csv")
```

```{r}
plot_eigen(waasb_model, size.lab = 14, size.tex.lab = 14)
```

The above output shows the eigenvalues and the proportion of variance explained by each principal component axis of the BLUP interaction effects matrix.

Phenotypic means

```{r}
data <- waasb_model$GY$MeansGxE
data
```

In this output, Y is the phenotypic mean for each genotype and environment combination (yij
), estimated by yij=∑kyij/B
 with k=1,2,...B
.

Biplots
Provided that an object of class waasb is available in the global environment, the graphics may be obtained using the function plot_scores(). To do that, we will revisit the previously fitted model WAASB . Please, refer to ?plot_scores for more details. Four types of graphics can be generated: 1 = PC1×PC2
; 2 = GY×PC1
; 3 = GY×WAASB
; and 4 = a graphic with nominal yield as a function of the environment PCA1 scores.

biplot type 1: GY x PC1

```{r}
c <- plot_scores(waasb_model, type = 1)
d <- plot_scores(waasb_model,
                 type = 1,
                 col.gen = "black",
                 col.env = "red",
                 col.segm.env = "red",
                 axis.expand = 1.5)
arrange_ggplot(c, d, tag_levels = list(c("c", "d")))
```
```{r}
d
```

biplot type 2: PC1 x PC2

```{r}

e <- plot_scores(waasb_model, type = 2)
f <- plot_scores(waasb_model,
                 type = 2,
                 polygon = TRUE,
                 col.segm.env = "transparent",
                 plot_theme = theme_metan_minimal())
arrange_ggplot(e, f, tag_levels = list(c("e", "f")))
```
```{r}
f
```
biplot type 3: GY x WAASB
The quadrants proposed by Olivoto et al. (2019) in the following biplot represent four classifications regarding the joint interpretation of mean performance and stability. The genotypes or environments included in quadrant I can be considered unstable genotypes or environments with high discrimination ability, and with productivity below the grand mean. In quadrant II are included unstable genotypes, although with productivity above the grand mean. The environments included in this quadrant deserve special attention since, in addition to providing high magnitudes of the response variable, they present a good discrimination ability. Genotypes within quadrant III have low productivity, but can be considered stable due to the lower values of WAASB. The lower this value, the more stable the genotype can be considered. The environments included in this quadrant can be considered as poorly productive and with low discrimination ability. The genotypes within the quadrant IV are highly productive and broadly adapted due to the high magnitude of the response variable and high stability performance (lower values of WAASB).

```{r}

g <- plot_scores(waasb_model, type = 3)
h <- plot_scores(waasb_model, type = 3,
                 x.lab = "My customized x label",
                 size.shape.gen = 3,
                 size.tex.gen = 2,
                 x.breaks = seq(1.5, 4.5, by = 0.5),
                 plot_theme = theme_metan(color.background = "white"))
arrange_ggplot(g, h, tag_levels = list(c("g", "h")))
```

```{r}
g
```
To obtain the WAASB index for a set of variables, the function get_model_data() is used, as shown bellow.

biplot type 4 : nominal yield and environment IPCA1

```{r}
i <- plot_scores(waasb_model, type = 4)
j <- plot_scores(waasb_model,
                 type = 4,
                 size.tex.gen = 1.5,
                 color = FALSE,
                 col.alpha.gen = 0,
                 col.alpha.env = 0,
                 plot_theme = theme_metan(color.background = "white"))
arrange_ggplot(i, j, tag_levels = list(c("i", "j")))
plot(i)
```

Simultaneous selection for mean performance and stability
The waasby index is used for genotype ranking considering both the stability (waasb) and mean performance (y) based on the following model (Olivoto et al. 2019).

waasbyi=(rYi×θY)+(rWi×θW)θY+θW

where waasbyi
 is the superiority index for the i-th genotype; rYi
 and rWi
 are the rescaled values (0-100) for the response variable (y) and the stability (WAAS or WAASB), respectively; θY
 and θW
 are the weights for mean performance and stability, respectively.

This index was also already computed and stored into AMMI_model>GY>model. An intuitively plot may be obtained by running

```{r}
i <- plot_waasby(waasb_model)
j <- plot_waasby(waasb_model, col.shape = c("gray20", "gray80"))
arrange_ggplot(i, j, tag_levels = list(c("e", "f")))
```

```{r}
i
```

In the following example, we will apply the function wsmp() to the previously fitted model waasb_model aiming at planning different scenarios of waasby estimation by changing the weights assigned to the stability and the mean performance.The number of scenarios is defined by the arguments increment. By default, twenty-one different scenarios are computed. In this case, the the superiority index waasby is computed considering the following weights: stability (waasb or waas) = 100; mean performance = 0. In other words, only stability is considered for genotype ranking. In the next iteration, the weights becomes 95/5 (since increment = 5). In the third scenario, the weights become 90/10, and so on up to these weights become 0/100. In the last iteration, the genotype ranking for WAASY or WAASBY matches perfectly with the ranks of the response variable.
Printing the model outputs

```{r}
waasb_model <- 
  waasb(yld,
        env = Environment,
        gen = Genotype,
        rep = Replication,
        resp = c(Severity,Chlorosis,SP,PY,Biomass,HI,SHW,GY,GGA,Chlorosis,a,u,b,v,CSI),
        random = "gen", #Default
        verbose = TRUE) #Default
```

```{r}
scenarios=wsmp(waasb_model)
```

In addition, the genotype ranking depending on the number of multiplicative terms used to estimate the WAAS index is also computed.

```{r}
scenarios$GY$hetdata
```

```{r}
scenarios$GY$hetdata
```

Plotting the heat map graphics
The first type of heatmap shows the genotype ranking depending on the number of principal component axes used for estimating the WAASB index. An euclidean distance-based dendrogram is used for grouping the genotypes based on their ranks. The second type of heatmap shows the genotype ranking depending on the WAASB/GY ratio. The ranks obtained with a ratio of 100/0 considers exclusively the stability for genotype ranking. On the other hand, a ratio of 0/100 considers exclusively the productivity for genotype ranking. Four clusters are estimated (1) unproductive and unstable genotypes; (2) productive, but unstable genotypes; (3) stable, but unproductive genotypes; and (4), productive and stable genotypes (Olivoto et al. 2019).

Ranks of genotypes depending on the number of PCA used to estimate the WAASB

```{r}

plot(scenarios, type = 1)
```

Ranks of genotypes depending on the WAASB/GY ratio

```{r}

plot(scenarios, type = 2)
```

Others BLUP-based stability indexes
Colombari Filho et al. (2013) have shown the use of three BLUP-based indexes for selecting genotypes with performance and stability. The first is the harmonic mean of genotypic values -or BLUPS- (HMGV) a stability index that considers the genotype with the highest harmonic mean across environments as the most stable, as follows:

HMGVi=E∑j=1E1BLUPij

The second is the relative performance of genotypic values (RPGV), an adaptability index estimated as follows:

RPGVi=1e∑j=1eBLUPij/μj

The third and last is the harmonic mean of relative performance of genotypic values (HMRPGV), a simultaneous selection index for stability, adaptability and mean performance, estimated as follows:

HMRPGVi=E∑j=1E1BLUPij/μj

```{r}
#library(dplyr)
#Res_ind <- 
#yld %>%
#gamem_met(GY,HI, PY, SP,PW, GW,SHW,Biomass, verbose = FALSE) %>% 
#blup_indexes()
# Warning: The WAASB index was not computed.
# Use an object computed with `waasb()` to get this index.
```

```{r}
model1=gamem_met(yld,
                 env = Environment,
                 gen = Genotype,
                 rep = Replication,
                 resp = c(Severity,Chlorosis,SP,PY,Biomass,HI,SHW,Necrosis,GGA,Chlorosis,a,u,GY,b,v,CSI))
```

```{r}
plot(model1,type = "re")
```

AI-BLUP selection index
The FAI-BLUP is a multi-trait index based on factor analysis and ideotype-design recentely proposed by Rocha, Machado, and Carneiro (2018). It is based on factor analysis, when the factorial scores of each ideotype are designed according to the desirable and undesirable factors. Then, a spatial probability is estimated based on genotype-ideotype distance, enabling genotype ranking. Here we will use the mixed-model mod as inpute data. By default, the selection is made to increase the value of all traits. Change this default with the arguments DI and UI.

```{r}
yld %>%
  gamem(Genotype, Replication, everything()) %>%
  fai_blup() %>%
  plot()
```

```{r}
yld %>%
  gamem(Genotype, Replication, resp = c(Severity,Chlorosis,SP,PY,Biomass,HI,SHW,GGA,Chlorosis,GY,b)) %>%
  fai_blup() %>%
  plot()
```

IMPORTANT fai_blup() recognizes models fitted with both gamem_met() and waasb(). For balanced data (all genotypes in all environments) waasb() and gamem_met() will return the same model. In case of unbalanced trials, the function waasb() will return an error since a complete two-way table is required to the singular value decomposition procedure.

Rendering engine
This vignette was built with pkgdown. All tables were produced with the package DT using the following function.

```{r}
library(DT) # Used to make the tables
# Function to make HTML tables
print_table <- function(table, rownames = FALSE, digits = 3, ...){
  df <- datatable(table, rownames = rownames, extensions = 'Buttons',
                  options = list(scrollX = TRUE, 
                                 dom = '<<t>Bp>',
                                 buttons = c('copy', 'excel', 'pdf', 'print')), ...)
  num_cols <- c(as.numeric(which(sapply(table, class) == "numeric")))
  if(length(num_cols) > 0){
    formatSignif(df, columns = num_cols, digits = digits)
  } else{
    df
  }
}
```


```{r}
a <- plot_scores(model1)
```

```{r}
b <- plot_scores(model1,
                 type = 2, # AMMI 2 biplot
                 polygon = TRUE, # show a polygon
                 highlight = c("Kasawaya", "Munzungu_munemari", "Chigunwe"), #highlight genotypes
                 col.alpha.env = 0.5, # alpha for environments
                 col.alpha.gen = 0, # remove the other genotypes
                 col.env = "gray", # color for environment point
                 col.segm.env = "gray", # color for environment segment
                 plot_theme = theme_metan_minimal()) # theme
arrange_ggplot(a, b, tag_levels = "a")
```

#####Computing parametric and non-parametric stability indexes
The easiest way to compute parametric and non-parametric stability indexes in metan is by using the function ge_stats(). It is a wrapper function around a lot of specific functions for stability indexes. To get the results into a “ready-to-read” file, use get_model_data() or its shortcut gmd()
```{r}
stats <- ge_stats(yld, Environment, Genotype, Farmer, GY)
stats=ge_stats(yld,env=Environment,gen = Genotype,rep = Farmer)
```

```{r}
get_model_data(stats)
```

```{r}
model <- gge(yld, Environment, Genotype, GY)
model2 <- gge(yld, Environment, Genotype, GY, svp = "genotype")
model3 <- gge(yld, Environment, Genotype, GY, svp = "symmetrical")
a <- plot(model)
b <- plot(model2, type = 8)
c <- plot(model2,
          type = 2,
          col.gen = "black",
          col.env = "gray70",
          axis.expand = 1.5,
          plot_theme = theme_metan_minimal())
arrange_ggplot(a, b, c, tag_levels = "a",ncol = 2)
```

```{r}
c
```

####Biplot type 2: Mean performance vs. stability
```{r}
gge_model <- gge(yld, Environment, Genotype, GY, svp = "genotype")
c <- plot(gge_model, type = 2)
d <- plot(gge_model,
          type = 2,
          col.gen = "black",
          col.env = "red",
          axis_expand = 1.5,
          plot_theme = theme_metan_minimal())
arrange_ggplot(c, d, tag_levels = list(c("c", "d")))
```

####Biplot type 3: Which-won-where

```{r}
gge_model <- gge(yld, Environment, Genotype, GY, svp = "symmetrical")
e <- plot(gge_model, type = 3)
f <- plot(gge_model,
          type = 3,
          size.shape.win = 5,
          large_label = 6,
          col.gen = "black",
          col.env = "gray",
          title = FALSE)
arrange_ggplot(e, f, tag_levels = list(c("e", "f")))
```

###Biplot type 4: Discriminativeness vs. representativeness

```{r}
g <- plot(gge_model, type = 4)
h <- plot(gge_model,
          type = 4,
          plot_theme = theme_metan_minimal())
arrange_ggplot(g, h, tag_levels = list(c("g", "h")))
```

###Biplot type 5: Examine an environment

```{r}
gge_model <- gge(yld, Environment, Genotype, GY, svp = "symmetrical")
i <- plot(gge_model, type = 5, sel_env = "E4")
j <- plot(gge_model,
          type = 5,
          sel_env = "E4",
          col.gen = "black",
          col.env = "black",
          size.text.env = 10,
          axis_expand = 1.5)
arrange_ggplot(i, j, tag_levels = list(c("i", "j")))
```

####Biplot type 6: Ranking environments

```{r}

gge_model <- gge(data_ge, ENV, GEN, GY)
k <- plot(gge_model, type = 6)
l <- plot(gge_model,
          type = 6,
          col.gen = "black",
          col.env = "black",
          col.circle = "red",
          col.alpha.circle = 0.5,
          size.text.env = 10,
          axis_expand = 1.5,
          plot_theme = theme_metan(color.background = "white"))
arrange_ggplot(k, l, tag_levels = list(c("k", "l")))
```

###Biplot type 7: Examine a genotype

```{r}
gge_model <- gge(yld, Environment, Genotype, GY, svp = "genotype")
m <- plot(gge_model, type = 7, sel_gen = "Njiva")
n <- plot(gge_model,
          type = 7,
          sel_gen = "Njiva",
          col.gen = "black",
          col.env = "black",
          size.text.env = 10,
          axis_expand = 1.5,
          plot_theme = theme_metan(grid = "both"))
arrange_ggplot(m, n, tag_levels = list(c("m", "n")))
```

###Biplot type 8: Ranking genotypes

```{r}
o <- plot(yld, type = 8)
p <- plot(gge_model,
          type = 8,
          col.gen = "black",
          col.env = "gray",
          size.text.gen = 6,
          plot_theme = theme_metan_minimal())
arrange_ggplot(o, p, tag_levels = list(c("o", "p")))
```

###Biplot type 9: Compare two genotypes

```{r}
gge_model <- gge(data_ge, ENV, GEN, GY, svp = "symmetrical")
q <- plot(gge_model, type = 9, sel_gen1 = "G8", sel_gen2 = "G10")
r <- plot(gge_model,
          type = 9,
          sel_gen1 = "G8",
          sel_gen2 = "G10",
          col.gen = "black",
          title = FALSE)
arrange_ggplot(q, r, tag_levels = list(c("q", "r")))
```

###Biplot type 10: Relationship among environments

```{r}
gge_model <- gge(data_ge, ENV, GEN, GY)
s <- plot(gge_model, type = 10)
t <- plot(gge_model,
          type = 10,
          col.gen = "black",
          title = FALSE)
arrange_ggplot(s, t, tag_levels = list(c("s", "t")))
```

#######Gamen
# run a mixed-effect model (genotype as random effect)
```{r}
mod <-
  gamem(yld,
        gen = Genotype,
        rep = Replication,
        resp = c(GY,Chlorosis,SP,PY,Biomass,HI,SHW,GGA,Chlorosis,Severity,b))
```
# BLUPs for genotypes
```{r}
gmd(mod, "blupg")
```
# Compute the MGIDI index
# Default options (all traits with positive desired gains)
# Equal weights for all traits
```{r}
mgidi_ind <- mgidi(mod)
```

```{r}
gmd(mgidi_ind, "MGIDI")
```
```{r}
write.csv(mgidi_ind$MGIDI,"MGIDI values for grin yield.csv")
```

# Higher weight for traits V1 and V4
# This will increase the probability of selecting H7 and H9
# 30% selection pressure
```{r}
mgidi_ind2 <-
   mgidi(mod,
         weights = c(1, .2, .2, 1),
         SI = 30)
gmd(mgidi_ind2, "MGIDI")
```

```{r}
write.csv(mgidi_ind2$MGIDI,"MGIDI 30 indexes for yyd.csv")
```

# plot the contribution of each factor on the MGIDI index
```{r}
p1 <- plot(mgidi_ind, type = "contribution")
p2 <- plot(mgidi_ind2, type = "contribution")
p1 + p2
```

# Positive desired gains for V1, V2 and V3
# Negative desired gains for V4
```{r}
mgidi_ind3 <-
  mgidi(mod,
       ideotype = c("h, h, h, h,h,l,l,l,l,l"))
```

####correlation analysis and heatmaps
read the data

```{r}
R1=read.csv("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/DPhil/RGB for indirect selection/Field trial data final/RGB_data_final/R1_means.csv")
attach(R1)
```

```{r}
library(pheatmap)
library(corrplot)
library(correlation)
library(RColorBrewer)
####BiocManager::install("ComplexHeatmap")
library(ComplexHeatmap)
```

```{r}
correlR1<-correlation(R1[,2:16],method="pearson")
```

```{r}
write.csv(correlR1,"Correlation coefficients for T File R1 variables.csv")
```

```{r}
corrR1=read.csv("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/DPhil/RGB for indirect selection/Field trial data final/RGB_data_final/Correlation coefficients for T File R1 variables.csv",row.names = "Trait")
attach(corrR1)
heat_plotR1=as.matrix(corrR1)
```

```{r}
P1 <-ComplexHeatmap::pheatmap(heat_plotR1, 
                      col = brewer.pal(10, 'RdYlGn'), # choose a colour scale for your data
                      cluster_rows = T, cluster_cols = T, # set to FALSE if you want to remove the dendograms
                      clustering_distance_cols = 'euclidean',
                      clustering_distance_rows = 'euclidean',
                      clustering_method = 'ward.D',
show_colnames = T, show_rownames = T,
display_numbers = TRUE,
         number_color = "black", 
         fontsize_number = 6,#
# displaying column and row names
                      main = "Flowering stage",
heatmap_legend_param = list(
        title = "Correlation", 
        at = c(-1, 0, 1), 
        labels = c("Negative", "zero", " Positive")))# a title for our heatmap
P1
```

```{r}
## Showing rows and columns
pheatmap(corrR1,
         main = "R2",
         show_colnames = T, show_rownames = T,
         number_color = "yellow", 
         fontsize_number = 8)
```

```{r}
## Showing values
pheatmap(corrR1,
         fontsize_col = 10,
         fontsize_row = 10,
         display_numbers = TRUE,
         number_color = "black", 
         fontsize_number = 6,#
         border_color = "black") # default is grey60
```

```{r}
R4=read.csv("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/DPhil/RGB for indirect selection/Field trial data final/RGB_data_final/R4_means.csv")
attach(R4)
```

```{r}
correlR4<-correlation(R4,method="pearson")
```

```{r}
write.csv(correlR4,"Correlation coefficient R4.csv")
```

```{r}
corrR4=read.csv("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/DPhil/RGB for indirect selection/Field trial data final/RGB_data_final/Correlation coefficient R4.csv",row.names = "Trait")
corrR4=as.matrix(corrR4)
```

```{r}
P4 <- ComplexHeatmap::pheatmap(corrR4, 
                      col = brewer.pal(10, 'RdYlGn'), # choose a colour scale for your data
                      cluster_rows = T, cluster_cols = T, # set to FALSE if you want to remove the dendograms
                      clustering_distance_cols = 'euclidean',
                      clustering_distance_rows = 'euclidean',
                      clustering_method = 'ward.D',
show_colnames = T, show_rownames = T,
display_numbers = TRUE,
         number_color = "black", 
         fontsize_number = 6,#
# displaying column and row names
                      main = "Pod maturity stage",
heatmap_legend_param = list(
        title = "Correlation", 
        at = c(-1, 0, 1), 
        labels = c("Negative", "zero", " Positive"))) # a title for our heatmap
P4
```

```{r}
## Showing rows and columns
pheatmap(corrRall,
         main = "Overall",
         show_colnames = T, show_rownames = T,
         number_color = "yellow", 
         fontsize_number = 8)
```

```{r}
## Showing values
pheatmap(corrR2,
         fontsize_col = 10,
         fontsize_row = 10,
         display_numbers = TRUE,
         number_color = "black", 
         fontsize_number = 6,#
         border_color = "black") # default is grey60
```
##############R6#################################################33
```{r}
R6=read.csv("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/DPhil/RGB for indirect selection/Field trial data final/RGB_data_final/R6_means.csv")
attach(R6)
```

```{r}
correlR6<-correlation(R6[,2:16],method="pearson")
```

```{r}
write.csv(correlR6,"Correlation coefficients of R6.csv")
```

```{r}
corrR6=read.csv("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/DPhil/RGB for indirect selection/Field trial data final/RGB_data_final/Correlation coefficients of R6.csv",row.names = "Trait")
attach(corrR6)
corrR7=as.matrix(corrR6)
```

```{r}
P6 <- pheatmap(corrR6, 
                      col = brewer.pal(10, 'RdYlGn'), # choose a colour scale for your data
                      cluster_rows = T, cluster_cols = T, # set to FALSE if you want to remove the dendograms
                      clustering_distance_cols = 'euclidean',
                      clustering_distance_rows = 'euclidean',
                      clustering_method = 'ward.D',
display_numbers = TRUE,
         number_color = "black", 
         fontsize_number = 7,#
# displaying column and row names
                      main = "Maturity stage",
heatmap_legend_param = list(
        title = "Correlation", 
        at = c(-1, 0, 1), 
        labels = c("Negative", "zero", " Positive"))) # a title for our heatmap
P6
```
#############Overal################################
```{r}
Rall=read.csv("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/DPhil/RGB for indirect selection/Field trial data final/RGB_data_final/overal_means.csv")
attach(Rall)
```

```{r}
correlRall<-correlation(Rall[,2:16],method="pearson")
```

```{r}
write.csv(correlRall,"Correlation coefficients of Rall.csv")
```

```{r}
corrRall=read.csv("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/DPhil/RGB for indirect selection/Field trial data final/RGB_data_final/Correlation coefficients of Rall.csv",row.names = "Trait")
attach(corrRall)
corrRall=as.matrix(corrRall)
```

```{r}
P4 <- pheatmap(corrRall, 
                      col = brewer.pal(10, 'RdYlGn'), # choose a colour scale for your data
                      cluster_rows = T, cluster_cols = T, # set to FALSE if you want to remove the dendograms
                      clustering_distance_cols = 'euclidean',
                      clustering_distance_rows = 'euclidean',
                      clustering_method = 'ward.D',
show_colnames = TRUE, show_rownames = TRUE,
display_numbers = TRUE,
         number_color = "black", 
         fontsize_number = 6,#
# displaying column and row names
                      main = "Overall",
heatmap_legend_param = list(
        title = "Correlation", 
        at = c(-1, 0, 1), 
        labels = c("Negative", "zero", " Positive"))) # a title for our heatmap
P4
```

```{r}
library(patchwork)
library(cowplot)
```

```{r}
citation(package = "randomForest")
```


