---
title: "Rust yield ANOVA error bar"
author: "Chinwa"
date: "2024-06-13"
output: html_document
---
####read the data

```{r}
setwd("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/DPhil/RGB for indirect selection/GLASSHOUSE FINAL EXPERIMENTS")
```

```{r}

```

```{r}
attach(GRV)
```

```{r}
library(ggplot2)
library(agricolae)
library(dplyr)
library(patchwork)
```
```{r}
attach(GRV)
```

```{r}
GRV$Genotypes=as.factor(GRV$Genotypes)
GRV$Replication=as.factor(GRV$Replication)
GRV$Group=as.factor(GRV$Group)
```

####Pod incubation period####
```{r}
value_max=GRV %>% group_by(Genotypes) %>% summarize(max_value=max(Incubation))
hsd=HSD.test(aov(Incubation~Genotypes+Replication,data=GRV),trt="Genotypes",group = T)
sig.letters=hsd$groups[order(row.names(hsd$groups)),]
```

```{r}
GRVinc=ggplot(data=GRV,aes(reorder(x=Genotypes,-Incubation),y=Incubation))+
  geom_boxplot(aes(fill=Genotypes))+
                 geom_text(data=value_max,aes(x=Genotypes,y=0.+
                                                            max_value,label=sig.letters$groups),vjust=0)+stat_boxplot(width=0.1)+ggtitle("")+xlab("")+ylab("Incubation period (days)")+theme_classic()+
  theme(axis.text.x = element_text(angle = 90,size = 10))+
  theme(axis.text.y = element_text(angle = 0,size = 15))+
  theme(axis.title.x = element_text(size = 15))+
  theme(axis.title.y = element_text(size = 15))+
  theme(legend.position = "none")
```

```{r}
GRVinc
```

####pod yield#####

```{r}
value_max=GRV %>% group_by(Genotypes) %>% summarize(max_value=max(GRV$Pod_yield))
hsd=HSD.test(aov(GRV$Pod_yield~Genotypes+Replication,data=GRV),trt="Genotypes",group = T)
sig.letters=hsd$groups[order(row.names(hsd$groups)),]
```

```{r}
GRVyld=ggplot(data=GRV,aes(reorder(x=Genotypes,-Pod_yield),y=Pod_yield))+
  geom_boxplot(aes(fill=Genotypes))+
                 geom_text(data=value_max,aes(x=Genotypes,y=0.5+
                                                            max_value,label=sig.letters$groups),vjust=0)+stat_boxplot(width=0.1
                                                                                                                      )+ggtitle("")+xlab("")+ylab("Pod yield (g/plant)")+theme_classic()+
  theme(axis.text.x = element_text(angle = 90,size = 10))+
  theme(axis.text.y = element_text(angle = 0,size = 15))+
  theme(axis.title.y = element_text(size = 15))+
  theme(axis.title.x = element_text(size = 10))+
  theme(legend.position = "none")
```

```{r}
GRVyld
```


####T file data####

```{r}
attach(GRV_T)
```
###subste the data growth stage 2
```{r}
##GRVT2=subset(GRV_T,Growth_stage=="R5",select = 1:18)
```

```{r}
##GRVT2
```

```{r}
GRV_T$Replication=as.factor(GRV_T$Rep)
GRV_T$Group=as.factor(GRV_T$Group)
GRV_T$Genotypes=as.factor(GRV_T$Genotype)
```

```{r}
value_max=GRV_T %>% group_by(Genotypes) %>% summarize(max_value=max(Severity))
hsd=HSD.test(aov(Severity~Genotypes+Replication,data=GRV_T),trt="Genotypes",group = T)
sig.letters=hsd$groups[order(row.names(hsd$groups)),] 
```

```{r}
GRVsev=ggplot(data=GRV_T,aes(x=reorder(Genotypes,-Severity),y=Severity))+
  geom_boxplot(aes(fill=Genotypes))+
                 geom_text(data=value_max,aes(x=Genotypes,y=0.1+
                                                            max_value,label=sig.letters$groups),vjust=0)+stat_boxplot(width=0.1)+ggtitle("")+xlab("")+ylab("GRV severity (%)")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90,size = 10))+
  theme(axis.text.y = element_text(angle = 0,size = 15))+
  theme(axis.title.y = element_text(size = 15))+
  theme(axis.title.x = element_text(size = 10))+
  theme(legend.position = "none")
```

```{r}
GRVsev
```

```{r}
value_max=GRV_T %>% group_by(Genotypes) %>% summarize(max_value=max(GGA))
hsd=HSD.test(aov(GGA~Genotypes+Replication,data=GRV_T),trt="Genotypes",group = T)
sig.letters=hsd$groups[order(row.names(hsd$groups)),]
```

```{r}
GRVGGA=ggplot(data=GRV_T,aes(reorder(x=Genotypes,-GGA),y=GGA))+
  geom_boxplot(aes(fill=Genotypes))+
                 geom_text(data=value_max,aes(x=Genotypes,y=0.1+
                                                            max_value,label=sig.letters$groups),vjust=0)+stat_boxplot(width=0.1)+ggtitle("")+xlab("")+ylab("GGA")+theme_classic()+
  theme(axis.text.x = element_text(angle = 90,size = 10))+
  theme(axis.text.y = element_text(angle = 0,size = 15))+
  theme(axis.title.y = element_text(size = 15))+
  theme(axis.title.x = element_text(size = 10))+
  theme(legend.position = "none")
```

```{r}
GRVGGA
```

```{r}
value_max=GRV_T %>% group_by(Genotypes) %>% summarize(max_value=max(GA))
hsd=HSD.test(aov(GA~Genotypes+Replication,data=GRV_T),trt="Genotypes",group = T)
sig.letters=hsd$groups[order(row.names(hsd$groups)),]
```

```{r}
GRVGA=ggplot(data=GRV_T,aes(reorder(x=Genotypes,-GA),y=GA))+
  geom_boxplot(aes(fill=Genotypes))+
                 geom_text(data=value_max,aes(x=Genotypes,y=0.1+
                                                            max_value,label=sig.letters$groups),vjust=0)+stat_boxplot(width=0.1)+ggtitle("")+xlab("")+ylab("GA")+theme_classic()+
  theme(axis.text.x = element_text(angle = 90,size = 10))+
  theme(axis.text.y = element_text(angle = 0,size = 15))+
  theme(axis.title.y = element_text(size = 15))+
  theme(axis.title.x = element_text(size = 10))+
  theme(legend.position = "none")
```

```{r}
GRVGA
```

```{r}
value_max=GRV_T %>% group_by(Genotypes) %>% summarize(max_value=max(a.))
hsd=HSD.test(aov(a.~Genotypes+Replication,data=GRV_T),trt="Genotypes",group = T)
sig.letters=hsd$groups[order(row.names(hsd$groups)),]
```

```{r}
GRVa=ggplot(data=GRV_T,aes(reorder(x=Genotypes,-a.),y=a.))+
  geom_boxplot(aes(fill=Genotypes))+
                 geom_text(data=value_max,aes(x=Genotypes,y=0.1+
                                                            max_value,label=sig.letters$groups),vjust=0)+stat_boxplot(width=0.1)+ggtitle("")+xlab("")+ylab("a*")+theme_classic()+
  theme(axis.text.x = element_text(angle = 90,size = 10))+
  theme(axis.text.y = element_text(angle = 0,size = 15))+
  theme(axis.title.y = element_text(size = 15))+
  theme(axis.title.x = element_text(size = 10))+ylim(-28,2)+
  theme(legend.position = "none")
```

```{r}
GRVa
```

```{r}
value_max=GRV_T %>% group_by(Genotypes) %>% summarize(max_value=max(CSI))
hsd=HSD.test(aov(CSI~Genotypes+Replication,data=GRV_T),trt="Genotypes",group = T)
sig.letters=hsd$groups[order(row.names(hsd$groups)),]
```

```{r}
GRVCSI=ggplot(data=GRV_T,aes(reorder(x=Genotypes,-CSI),y=CSI))+
  geom_boxplot(aes(fill=Genotypes))+
                 geom_text(data=value_max,aes(x=Genotypes,y=0.1+
                                                            max_value,label=sig.letters$groups),vjust=0)+stat_boxplot(width=0.1)+ggtitle("")+xlab("")+ylab("CSI")+theme_classic()+
  theme(axis.text.x = element_text(angle = 90,size = 10))+
  theme(axis.text.y = element_text(angle = 0,size = 15))+
  theme(axis.title.y = element_text(size = 15))+
  theme(axis.title.x = element_text(size = 10))+
  theme(legend.position = "none")
```

```{r}
GRVCSI
```

########################leafspots########################################

```{r}
LSP=read.csv("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/DPhil/RGB for indirect selection/GLASSHOUSE FINAL EXPERIMENTS/leafspot_yield.csv")
```

```{r}
attach(LSP)
```

```{r}
LSP$Genotype=as.factor(LSP$Genotype)
LSP$Genotype=as.factor(LSP$Genotype)
LSP$Replication=as.factor(LSP$Replication)
LSP$Group=as.factor(LSP$Group)
```

```{r}
library(agricolae)
```

```{r}
attach(LSP)
```

```{r}
model<-aov(Pod_yield~Genotype, data=LSP)
out <- LSD.test(model,"Genotype", group=TRUE,console=TRUE,
main="fun")
```

#stargraph
# Variation range: max and min

```{r}
plot(out)
```
#endgraph

```{r}
value_max2=LSP %>% group_by(Genotype) %>% summarize(max_value2=max(Pod_yield))
hsd=LSD.test(aov(Pod_yield~Genotype,data=LSP),trt="Genotype",group = T)
sig.letters=hsd$groups[order(row.names(hsd$groups)),]
```

```{r}
LSPyld=ggplot(data=LSP,aes(reorder(x=Genotype,-Pod_yield),y=Pod_yield))+
  geom_boxplot(aes(fill=Genotypes))+
                 geom_text(data=value_max2,aes(x=Genotype,y=0.1+
                                                            max_value2,label=sig.letters$groups),vjust=0)+stat_boxplot(width=0.1)+ggtitle("")+xlab("")+ylab("Pod yield (g/plant)")+theme_classic()+
  theme(axis.text.x = element_text(angle = 90,size = 10))+
theme(axis.text.y = element_text(angle = 0,size = 15))+  
  theme(axis.title.x = element_text(size = 10))+
  theme(axis.title.x = element_text(size = 10))+
  theme(axis.title.y = element_text(size = 15))+
  theme(legend.position = "none")
```

```{r}
LSPyld
```
############Incubation period#############################

```{r}
value_max2=LSP %>% group_by(Genotype) %>% summarize(max_value2=max(Incubation))
hsd=LSD.test(aov(Incubation~Genotype+Replication,data=LSP),trt="Genotype",group = T)
sig.letters=hsd$groups[order(row.names(hsd$groups)),]
```

```{r}
LSPinc=ggplot(data=LSP,aes(reorder(x=Genotype,-Incubation),y=Incubation))+
  geom_boxplot(aes(fill=Genotype))+
                 geom_text(data=value_max2,aes(x=Genotype,y=0.3+
                                                            max_value2,label=sig.letters$groups),vjust=0)+stat_boxplot(width=0.1)+ggtitle("")+xlab("")+ylab("Incubation period (days)")+theme_classic()+
  theme(axis.text.x = element_text(angle = 90,size = 10))+
  theme(axis.text.y = element_text(angle = 0,size = 15))+
  theme(axis.title.x = element_text(size = 10))+
  theme(axis.title.y = element_text(size = 15))+
  theme(legend.position = "none")
```

```{r}
LSPinc
```

```{r}
attach(LSPot)
```
```{r}
LSPot$Stage=as.factor(LSPot$Stage)
LSPot$Replication=as.factor(LSPot$Replication)
LSPot$Groups=as.factor(LSPot$Groups)
LSPot$Genotypes=as.factor(LSPot$Genotypes)
```

```{r}
value_max2=LSPot %>% group_by(Genotypes) %>% summarize(max_value2=max(GGA))
hsd=LSD.test(aov(GGA~Genotypes+Replication,data=LSPot),trt="Genotypes",group = T)
sig.letters=hsd$groups[order(row.names(hsd$groups)),]
```

```{r}
LSPotGGA=ggplot(data=LSPot,aes(reorder(x=Genotypes,-GGA),y=GGA))+
  geom_boxplot(aes(fill=Genotypes))+
                 geom_text(data=value_max2,aes(x=Genotypes,y=0.1+
                                                            max_value2,label=sig.letters$groups),vjust=0)+stat_boxplot(width=0.1)+ggtitle("")+xlab("")+ylab("GGA")+theme_classic()+
  theme(axis.text.x = element_text(angle = 90,size = 10))+
theme(axis.text.y = element_text(angle = 0,size = 15))+
  theme(axis.title.x = element_text(size = 15))+
  theme(axis.title.y = element_text(size = 15))+
  theme(legend.position = "none")
```

#####Severity score###################

```{r}
LSPT=read.csv("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/DPhil/RGB for indirect selection/GLASSHOUSE FINAL EXPERIMENTS/leafspot.csv")
```

```{r}
attach(LSPT)
```
```{r}
LSPT$Genotype=as.factor(LSPT$Genotype)
LSPT$Replication=as.factor(LSPT$Replication)
```

```{r}
value_max2=LSPT %>% group_by(Genotype) %>% summarize(max_value2=max(GA))
hsd=LSD.test(aov(GA~Genotype,data=LSPT),trt="Genotype",group = T)
sig.letters=hsd$groups[order(row.names(hsd$groups)),]
```

```{r}
LSPT1=ggplot(data=LSPT,aes(reorder(x=Genotype,-GA),y=GA))+
  geom_boxplot(aes(fill=Genotype))+
                 geom_text(data=value_max2,aes(x=Genotype,y=0+
                                                            max_value2,label=sig.letters$groups),vjust=0)+stat_boxplot(width=0.1)+ggtitle("")+xlab("")+ylab("GA)")+theme_classic()+
  theme(axis.text.x = element_text(angle = 90,size = 10))+
  theme(axis.text.y = element_text(angle = 0,size = 15))+
  theme(axis.title.x = element_text(size = 15))+
  theme(axis.title.y = element_text(size = 15))+
  theme(legend.position = "none")
```

```{r}
LSPT1
```


###########################R7###############################################


#####################GGA###################################################

```{r}
R7=read.csv("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/DPhil/RGB for indirect selection/GLASSHOUSE FINAL EXPERIMENTS/R7.csv")
```

```{r}
model<-aov(v.~Genotype, data=R7)
out <- LSD.test(model,"Genotype", group=TRUE,console=TRUE,
main="fun")
```
```{r}
attach(R7)
```

```{r}
R7$Genotype=as.factor(R7$Genotype)
R7$Replication=as.factor(R7$Replication)
```

```{r}
str(R7)
```

```{r}
value_maxR7=R7 %>% group_by(Genotype) %>% summarize(max_value2=max(CSI))
hsd=LSD.test(aov(CSI~Genotype,data=R7),trt="Genotype",group = T)
sig.letters=hsd$groups[order(row.names(hsd$groups)),]
```

```{r}
R7GGC=ggplot(data=R7,aes(reorder(x=Genotype,-CSI),y=CSI))+
  geom_boxplot(aes(fill=Genotype))+
                 geom_text(data=value_maxR7,aes(x=Genotype,y=0.1+
                                                            max_value2,label=sig.letters$groups),vjust=0)+stat_boxplot(width=0.1)+ggtitle("")+xlab("Genotype")+ylab("CSI (%)")+theme_classic()+
  theme(axis.text.x = element_text(angle = 90,size = 10))+
theme(axis.text.y = element_text(angle = 0,size = 15))+
  theme(axis.title.x = element_text(size = 15))+
  theme(axis.title.y = element_text(size = 15))+
  theme(legend.position = "none")
```

```{r}
R7GGC
```





```{r}
value_max2=LSPT %>% group_by(Genotype) %>% summarize(max_value2=max(v.))
hsd=HSD.test(aov(v.~Genotype+Replication,data=LSPT),trt="Genotype",group = T)
sig.letters=hsd$groups[order(row.names(hsd$groups)),]
```

```{r}
LSPTGA=ggplot(data=LSPT,aes(reorder(x=Genotype,-v.),y=v.))+
  geom_boxplot(aes(fill=Genotype))+
                 geom_text(data=value_max2,aes(x=Genotype,y=0.1+
                                                            max_value2,label=sig.letters$groups),vjust=0)+stat_boxplot(width=0.1)+ggtitle("")+xlab("")+ylab("GA")+theme_classic()+
  theme(axis.text.x = element_text(angle = 90,size = 10))+
theme(axis.text.y = element_text(angle = 0,size = 15))+
  theme(axis.title.x = element_text(size = 10))+
  theme(axis.title.y = element_text(size = 15))+
  theme(legend.position = "none")
```

```{r}
LSPTGA
```

############################a.################################
```{r}
value_max2=LSPot %>% group_by(Genotypes) %>% summarize(max_value2=max(a.))
hsd=HSD.test(aov(a.~Genotypes+Replication,data=LSPot),trt="Genotypes",group = T)
sig.letters=hsd$groups[order(row.names(hsd$groups)),]
```

```{r}
LSPota.=ggplot(data=LSPot,aes(reorder(x=Genotypes,-a.),y=a.))+
  geom_boxplot(aes(fill=Genotypes))+
                 geom_text(data=value_max2,aes(x=Genotypes,y=0.1+
                                                            max_value2,label=sig.letters$groups),vjust=0)+stat_boxplot(width=0.1)+ggtitle("")+xlab("")+ylab("a*")+theme_classic()+
  theme(axis.text.x = element_text(angle = 90,size = 10))+
theme(axis.text.y = element_text(angle = 0,size = 10))+
  theme(axis.title.x = element_text(size = 10))+
  theme(axis.title.y = element_text(size = 15))+
  theme(legend.position = "none")
```

```{r}
LSPota.
```

```{r}
value_max2=LSPot %>% group_by(Genotypes) %>% summarize(max_value2=max(CSI))
hsd=HSD.test(aov(CSI~Genotypes+Replication,data=LSPot),trt="Genotypes",group = T)
sig.letters=hsd$groups[order(row.names(hsd$groups)),]
```

```{r}
LSPotCSI=ggplot(data=LSPot,aes(reorder(x=Genotypes,-CSI),y=CSI))+
  geom_boxplot(aes(fill=Genotypes))+
                 geom_text(data=value_max2,aes(x=Genotypes,y=0.1+
                                                            max_value2,label=sig.letters$groups),vjust=0)+stat_boxplot(width=0.1)+ggtitle("")+xlab("")+ylab("CSI")+theme_classic()+
  theme(axis.text.x = element_text(angle = 90,size = 10))+
theme(axis.text.y = element_text(angle = 0,size = 15))+
  theme(axis.title.x = element_text(size = 10))+
  theme(axis.title.y = element_text(size = 15))+
  theme(legend.position = "none")
```

```{r}
LSPotCSI
```
```{r}
library(patchwork)
library(ggpubr)
library(gridExtra)
```

```{r}
ggsave("LSPotCSI.JPG",width =8,height = 10,dpi = 300 )
```

####correlation analysis and heatmaps
read the data

```{r}
R1=read.csv("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/DPhil/RGB for indirect selection/GLASSHOUSE FINAL EXPERIMENTS/R1_means.csv")
attach(R1)
```

```{r}
library(pheatmap)
library(corrplot)
library(correlation)
library(RColorBrewer)
####BiocManager::install("ComplexHeatmap")
library(ComplexHeatmap)
```

```{r}
correlR1<-correlation(R1[,2:13],method="pearson")
```

```{r}
write.csv(correlR1,"Correlation coefficients for T File R1 variables.csv")
```

```{r}
corrR1=read.csv("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/DPhil/RGB for indirect selection/GLASSHOUSE FINAL EXPERIMENTS/Correlation coefficients for T File R1 variables.csv",row.names = "Trait")
attach(corrR1)
```

```{r}
heat_plotR2=as.matrix(corrR1)
```

```{r}
P1 <-ComplexHeatmap::pheatmap(heat_plotR2, 
                      col = brewer.pal(10, 'RdYlGn'), # choose a colour scale for your data
                      cluster_rows = T, cluster_cols = T, # set to FALSE if you want to remove the dendograms
                      clustering_distance_cols = 'euclidean',
                      clustering_distance_rows = 'euclidean',
                      clustering_method = 'ward.D',
show_colnames = T, show_rownames = T,
display_numbers = TRUE,
         number_color = "black", 
         fontsize_number = 6,#
# displaying column and row names
                      main = "R1",
heatmap_legend_param = list(
        title = "Correlation", 
        at = c(-1, 0, 1), 
        labels = c("Negative", "zero", " Positive")))# a title for our heatmap
P1
```

```{r}
## Showing rows and columns
pheatmap(corrR1,
         main = "R1",
         
         show_colnames = T, show_rownames = T,
         number_color = "yellow", 
         fontsize_number = 8)
```

```{r}
## Showing values
pheatmap(corrR1,
         fontsize_col = 10,
         fontsize_row = 10,
         display_numbers = TRUE,
         number_color = "black", 
         fontsize_number = 6,#
         border_color = "black") # default is grey60
```

```{r}
R4=read.csv("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/DPhil/RGB for indirect selection/GLASSHOUSE FINAL EXPERIMENTS/R4_means.csv")
attach(R4)
```

```{r}
correlR4<-correlation(R4[,2:13],method="pearson")
```

```{r}
write.csv(correlR4,"Correlation coefficient R4.csv")
```

```{r}
corrR4=read.csv("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/DPhil/RGB for indirect selection/GLASSHOUSE FINAL EXPERIMENTS/Correlation coefficient R4.csv",row.names = "Trait")
corrR4=as.matrix(corrR4)
```

```{r}
P2 <- ComplexHeatmap::pheatmap(corrR4, 
                      col = brewer.pal(10, 'RdYlGn'), # choose a colour scale for your data
                      cluster_rows = T, cluster_cols = T, # set to FALSE if you want to remove the dendograms
                      clustering_distance_cols = 'euclidean',
                      clustering_distance_rows = 'euclidean',
                      clustering_method = 'ward.D',
show_colnames = T, show_rownames = T,
display_numbers = TRUE,
         number_color = "black", 
         fontsize_number = 6,#
# displaying column and row names
                      main = "R4",
heatmap_legend_param = list(
        title = "Correlation", 
        at = c(-1, 0, 1), 
        labels = c("Negative", "zero", " Positive"))) # a title for our heatmap
P2
```

```{r}
## Showing rows and columns
pheatmap(corrR4,
         main = "Overall",
         show_colnames = T, show_rownames = T,
         number_color = "yellow", 
         fontsize_number = 8)
```

```{r}
## Showing values
pheatmap(corrR4,
         fontsize_col = 10,
         fontsize_row = 10,
         display_numbers = TRUE,
         number_color = "black", 
         fontsize_number = 6,#
         border_color = "black") # default is grey60
```
##############R3#################################################33
```{r}
R7=read.csv("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/DPhil/RGB for indirect selection/GLASSHOUSE FINAL EXPERIMENTS/R7_means.csv")
attach(R7)
```

```{r}
correlR7<-correlation(R7[,2:13],method="pearson")
```

```{r}
write.csv(correlR7,"Correlation coefficients of R7.csv")
```

```{r}
corrR7=read.csv("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/DPhil/RGB for indirect selection/GLASSHOUSE FINAL EXPERIMENTS/Correlation coefficients of R7.csv",row.names = "Trait")
attach(corrR7)
corrR7=as.matrix(corrR7)
```

```{r}
P3 <- pheatmap(corrR7, 
                      col = brewer.pal(10, 'RdYlGn'), # choose a colour scale for your data
                      cluster_rows = T, cluster_cols = T, # set to FALSE if you want to remove the dendograms
                      clustering_distance_cols = 'euclidean',
                      clustering_distance_rows = 'euclidean',
                      clustering_method = 'ward.D',
display_numbers = TRUE,
         number_color = "black", 
         fontsize_number = 6,#
# displaying column and row names
                      main = "R7",
heatmap_legend_param = list(
        title = "Correlation", 
        at = c(-1, 0, 1), 
        labels = c("Negative", "zero", " Positive"))) # a title for our heatmap
P3
```
#############Overal################################
```{r}
Rall=read.csv("E:/Field trials/Groundnut/data/Field trial data final/means_A_file_all.csv")
attach(Rall)
```

```{r}
correlRall<-correlation(Rall[,2:22],method="pearson")
```

```{r}
write.csv(correlRall,"Correlation coefficients of Rall.csv")
```

```{r}
corrRall=read.csv("E:/Field trials/Groundnut/data/Field trial data final/Correlation coefficients of Rall.csv",row.names = "Trait")
attach(corrRall)
corrRall=as.matrix(corrRall)
```

```{r}
P4 <- pheatmap(corrRall, 
                      col = brewer.pal(10, 'RdYlGn'), # choose a colour scale for your data
                      cluster_rows = T, cluster_cols = T, # set to FALSE if you want to remove the dendograms
                      clustering_distance_cols = 'euclidean',
                      clustering_distance_rows = 'euclidean',
                      clustering_method = 'ward.D',
show_colnames = TRUE, show_rownames = TRUE,
display_numbers = TRUE,
         number_color = "black", 
         fontsize_number = 6,#
# displaying column and row names
                      main = "Overall",
heatmap_legend_param = list(
        title = "Correlation", 
        at = c(-1, 0, 1), 
        labels = c("Negative", "zero", " Positive"))) # a title for our heatmap
P4
```
```{r}
library(patchwork)
library(cowplot)
```

#######principal component analysis###############################


```{r}
means=read.csv("C:/Users/hchin/OneDrive/Atlantic Beach/Documents/PhD/DPhil/DPhil/RGB for indirect selection/GLASSHOUSE FINAL EXPERIMENTS/A_file_all.csv",header = T)
attach(means)
```

```{r}
means$Genotype=as.factor(means$Genotype)
```

```{r}
library(FactoMineR)
library(factoextra)
library(ggplot2)
library(ggpubr)
```

```{r}
df=means[,-1]
row.names(df)=as.factor(`Genotype`)
```

```{r}
View(df)
```

```{r}
res.PCA <- PCA(df, graph = TRUE)
```

```{r}
print(res.PCA)
```

```{r}
eig.val <- get_eigenvalue(res.PCA)
```

```{r}
eig.val
```

```{r}
head(eig.val)
```

```{r}
eig.val=as.data.frame(eig.val)
```

```{r}
eigenvals=write.csv(eig.val,"Eigen values for the treatments.csv")
```

```{r}
write_xls(eig.val,"Eigen values for the parameters.xlsx")
```

```{r}
fviz_screeplot(res.PCA)
```

```{r}
var <- get_pca_var(res.PCA)
```

```{r}
var$coord
var$cor
var$contrib
```

```{r}
var$coord=as.data.frame(var$coord)
```

```{r}
var$cor=as.data.frame(var$cor)
```

```{r}
var$contrib=as.data.frame(var$contrib)
```

```{r}
write_xlsx(var$coord,"Coordinates for the parameters.xlsx")
```

```{r}
write_xlsx(var$cor,"correlations of the parameters .xlsx")
```

```{r}
write_xlsx(var$contrib,"contribution variances for the parameters.xlsx")
```

```{r}
write.table(var$coord,"Coord of variables.xls", row.names=TRUE, sep="\t")
```

```{r}
write.table(var$cor,"Correlation_variables_dimensions.xls", row.names=TRUE, sep="\t")
```

```{r}
write.table(var$contrib,"Contribution_variables.xls", row.names=TRUE, sep="\t")
```

# Plot of variables
```{r}
fviz_pca_var(res.PCA, repel = TRUE)
```
# Contribution to the first dimension
```{r}
a1=fviz_contrib(res.PCA, "var", axes = 1)
a1
```
# Contribution to the second dimension
```{r}
b=fviz_contrib(res.PCA, "var", axes = 2)
b
```

```{r}
c=fviz_contrib(res.PCA, "var", axes = 2)
c
```

```{r}
ggarrange(a1,b,c,ncol=1,nrow = 2)
```

# Compute hierarchical clustering on principal components
```{r}
res.hcpc <- HCPC(res.PCA, graph = FALSE)
res.hcpc
```

```{r}
varA=fviz_pca_var(res.PCA,axes = c(1, 2),col.var = "cos2",
                  gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
                  repel = TRUE )+
  theme_minimal()
varA
```

```{r}
varB=fviz_pca_var(res.PCA,axes = c(1, 2),col.var = "cos2",
                  gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
                  repel = TRUE )+
  theme_minimal()
```

```{r}

varC=fviz_pca_var(res.PCA,axes = c(1, 2),col.var = "cos2",
                  gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
                  repel = TRUE )+
  theme_minimal()
```

```{r}
ggarrange(varA,varB,ncol=2,legend = NULL,common.legend = TRUE)
```

```{r}
c=fviz_dend(res.hcpc,
            cex = 0.8,                   # Label size
            # Color palette see ?ggpubr::ggpar
            rect = FALSE, rect_fill = FALSE, # Add rectangle around groups
            horiz=TRUE, # Rectangle color
            labels_track_height = 0.5       # Augment the room for labels
)
c
```

```{r}
d=fviz_cluster(res.hcpc,axes = c(1, 2),
               repel = TRUE,           # Avoid label overlapping
               show.clust.cent = FALSE,geom=c("point","text"), # Show cluster centers
               # Color palette see ?ggpubr::ggpar
               ggtheme = theme_minimal(),ellipse.type = "convex",
               main = "Factor map"
)
d
```


```{r}
res.hcpc
clustdata=res.hcpc$data.clust
```

```{r}
write.table(  clustdata,"Dataset with the cluster of the individuals.xls", row.names=TRUE, sep="\t")
```

```{r}
quanti=res.hcpc$desc.var$quanti
quanti$`1`
quanti$`2`
quanti$`3`
```

```{r}
write.table(  quanti$`1`,"Description of the cluster 1 by the var.xls", row.names=TRUE, sep="\t")
```

```{r}
write.table(  quanti$`2`,"Description of the cluster 2 by the var.xls", row.names=TRUE, sep="\t")
```

```{r}
write.table(  quanti$`3`,"Description of the cluster 3 by the var.xls", row.names=TRUE, sep="\t")
```

```{r}
res.hcpc$desc.axes$quanti
```

```{r}
write.table(  res.hcpc$desc.axes$quanti,"Description of the clusters by the axes.xls", row.names=TRUE, sep="\t")
```

